Death By Dropdown?

### 2025-04-09 conference-materials/2025_shiny-conf/slides.qmd
Refines slides for clarity and conciseness
- Reorganizes {bidux} code sections and updates wording for better flow and engagement.
- Revises notes to be more concise

- **Stage 1 (Notice):** Users were overwhelmed with options
  * This relates to 'cognitive load' - the mental effort needed to process information
  * When the brain gets overtaxed, comprehension drops
  * Hick's Law shows decision time increases logarithmically with more choices
- **Stage 2 (Interpret):** We identified what users needed - increased Processing Fluency
  * When information is easy to process, people find it more credible
  * We focused on the key story executives needed, not showing all possible data
- **Stage 3 (Structure):** Moving to the Structuring stage...applied the 'Principle of Proximity' - which is a fancy way of saying we started grouping related elements together. 
  * Dual-Processing Theory distinguishes between fast, intuitive thinking and slower, deliberative thinking - we designed for both
  * We created logical groupings of metrics for natural data flow
- Now let's look at the code implementation
:::

@@

- For the later BID stages, we addressed subtler psychological factors affecting interpretation and action
- **Stage 4 (Anticipate):** We expected some folks to make flawed comparisons if they didn't have proper context.
  * This relates to the 'Anchoring Effect' - where people rely too heavily on the first piece of information they see
  * We also saw teams struggling to interpret the data in the same way, so we addressed this with something called a "framing toggle"...more on that in just a moment
- **Stage 5 (Validate):** Finally, we applied the 'Peak-End Rule' - the idea that people judge experiences by their most intense moment and ending
  * We created summary cards highlighting key takeaways so users leave with a clear, memorable impression
:::

@@

@@

::: {.r-stack}
::: {.fragment .fade-out}
```{r}
#| echo: true
library(bidux)

@@

@@

# Concept dictionary with fuzzy searching and case-insensitive lookup
viz_concepts <- bid_concepts("vis") 
paste("Found", nrow(viz_concepts), "visualization-related concepts")
```
:::
:::

# BID Notice stage
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  theory = "cognitive load",
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

@@

- "Let's look at what {bidux} offers today"
- "The concept dictionary provides implementation advice, not just theory"
- "You can search across psychology, UI/UX, and data visualization fields...and more!"
- "The workflow functions guide you through each BID stage"
- "Auto-suggestion of theories based on your problem description"
- "Component suggestions provide concrete UI implementation ideas"
- When the link appears: "We're launching this at ShinyConf today - check out the repo for full documentation"
:::

@@

- The results speak for themselves
- While quantitative metrics are still being collected, qualitative feedback has been overwhelmingly positive
- Users who previously abandoned dashboards now rely on them daily
- Decisions that took days now happen in minutes
- When testimonial appears: The emerging message is clear: this framework transforms how teams interact with their data
- "Remember, this wasn't just about visual improvements - this was a fundamental shift in how our users...and developers...approached the dashboard"
:::

@@

- "BID reduces friction between users and insights by applying psychological principles in a systematic way"
- "The tools we've shown today - reactable and echarts4r - give you everything you need to get started"
- "The bidux package makes these principles accessible to every R developer, guiding you through each stage"
- "Start with small implementations of these principles - even one or two can dramatically improve your dashboard"
- "Every dropdown you eliminate or visual cue you add helps someone make a better decision"
- "This is about making your users successful, not just making dashboards"
:::

>>>>

- **Stage 1 (Notice):** Users were overwhelmed with options
  * This relates to 'cognitive load' - the mental effort needed to process information
  * When the brain gets overtaxed, comprehension drops
  * Hick's Law shows decision time increases logarithmically with more choices
- **Stage 2 (Interpret):** We identified what users needed - increased Processing Fluency
  * When information is easy to process, people find it more credible
  * We focused on the key story executives needed, not showing all possible data
- **Stage 3 (Structure):** We applied the 'Principle of Proximity' in the structure stage...which is about grouping related elements together
  * We also make use of Dual-Processing Theory (which distinguishes between fast, intuitive thinking and slower, deliberative thinking) in the design to create logical groupings of metrics for more intuitive workflows
- Now let's see how we brought these concepts to life
:::

@@

- For the later BID stages, we addressed subtler psychological factors affecting interpretation and action
- **Stage 4 (Anticipate):** We expected some folks to make flawed comparisons without proper context
  * This relates to the 'Anchoring Effect'...where people rely too heavily on the first piece of information they see
  * We also saw teams struggling to interpret the data in the same way, which connects to 'Framing & Loss Aversion'...how the presentation of identical data as a "gain" or a "loss" significantly affects decisions
- **Stage 5 (Validate):** Finally, we applied the 'Peak-End Rule'...the idea that people judge experiences by their most intense moment and ending
  * This helped us understand why users needed clear, memorable takeaways to conclude their dashboard experience
- These psychological principles guided our approach to the remaining implementation
:::

@@

::: {.fragment}
```{r}
#| echo: true
library(bidux)

@@

@@

# Concept dictionary with fuzzy searching and case-insensitive lookup
cog_concepts <- bid_concepts("cog")
paste("Found", nrow(cog_concepts), "cog-related concepts")
```
:::

@@

# BID Notice stage targeting cognitive load theory
notice_stage <- bid_notice(
  problem = "Users overwhelmed by filtering options and dropdown menus",
  theory = "Cognitive Load Theory",
  evidence = "Feedback shows 65% of users abandon the dashboard after first use"
)

@@

- Currently, `{bidux}` has helper functions for all 5 stages of the BID framework [CLICK]
- **Notice tab**
  * `bid_notice` let's you begin by documenting user problems
  * Here we document user problems like "overwhelming dropdown menus" leading to high dashboard abandonment
  * Without specifying a theory, the package suggests one and provides actionable recommendations [SWITCH ON PRESENTATION SCREEN]
- **Concepts tab**
  * If you're unfamiliar with the concept or theory, you can use `bid_concepts` to look them up...[CLICK]...[CLICK]...[CLICK]...or search broadly (like "cog" for cognitive concepts)
  * Maybe you want to update your project to use a difference concept (e.g, cognitive load theory) [CLICK]
  * If so, you can go back and apply them directly in your workflow for new recommendations [SWITCH ON PRESENTATION SCREEN]
- **Chained Stages tab**
  * The real power though is piping through the entire framework
  * We start with our noticed problem, interpret the user's needs, then structure the solution...[CLICK]...[CLICK]...all while getting tailored suggestions at each stage [SWITCH ON PRESENTATION SCREEN]
- **Components tab**
  * And finally, depending on the unique needs of your project, `bid_suggest_components` bridges theory to code with UI component recommendations and implementation examples...[CLICK]...[CLICK]
- This is Phase 1 we're launching at ShinyConf today...[CLICK]...check our GitHub repo for updates
:::

@@

- The results speak for themselves
- While quantitative metrics are still being collected, qualitative feedback has been overwhelmingly positive
- Users who previously abandoned dashboards now rely on them daily
- Decisions that took days now happen in minutes
- The emerging message is clear: this framework transforms how teams interact with their data
- "Remember, this wasn't just about visual improvements - this was a fundamental shift in how our users...and developers...approached the dashboard"
:::

@@

- "BID reduces friction between users and insights by applying psychological principles in a systematic way"
- "The tools we've shown today - reactable and echarts4r - give you everything you need to get started"
- "The bidux package makes these principles accessible to every R developer, guiding you through each stage"
- Build dashboards that guide — not fry — your users
- "Every dropdown you eliminate or visual cue you add helps someone make a better decision"
:::
